Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures

### 2012-02-08 ltx/paper_short_2.tex
I really need to clean up this repo. sorry. the  main tex file is paper_short_2.tex

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

\section{Introduction}

We define the task of automated detection and tracking of arbitrary objects in videos to \willie{consist of three actions} be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking'). An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization. Algorithms able to be applied to a variety of objects and video types are particularly desirable. This paper \willie{introduces a new framework involving the use of dependent Dirichlet process mixture models. This framework provides the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.} 
% serves to introduce a new technique involving the use of dependent Dirichlet process mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking of multiple objects while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to perform this extraction and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.

% related work
% ------------

\subsection{Related Work}
\label{sec:relatedwork}

Related work involves studies related to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.

Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

Localization is related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving model-based clustering of individual pixels, local image texture representations, or other patterns found in image regions \cite{jain1997object, fei2005bayesian, sivic2005discovering}. These works usually aim to automatically segment images into meaningful regions. Note that the popular phrase `object detection' is sometimes used to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection, which incorporates both the task of determining what constitutes distinct objects and localizing these objects), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. An overview of object tracking methods can be found in \cite{yilmaz2006object}. Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges to a number of modes. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks. The technique we introduce in this paper falls into this category of methods, as it aims to provide a general unsupervised detection and tracking method for arbitrary objects in videos via clustering. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.

% extra background that I decided to cut (for now)
% ------------------------------------------------

% Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.

% Extraction
%

% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked. This method is beneficial because it optimizes the search for the `next' position of an object (i.e. the search for the region in a frame which is most similar to a target region in the previous frame). This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}. Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features. Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position. However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search. Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels. Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.

% Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}. Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .

% more background that I decided to cut (for now)
% -----------------------------------------------

% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).

% % Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.

% % extensions to multiple visual object tracking. problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
% % Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
% % End with description of initialization problem.

% % Methods for detection and tracking of objects. Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene). Supervised learning of specific targets to allow for initialization and provide better localization throughout. Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods. Particle filters with integrated detection (at least two of these). Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]). The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data. Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.

% % When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects. Two classic algorithms for MTT are known as MHT and JPDA. In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking). MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research. In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).

% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking). Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking. Work of Pece, MTT of multiple targets work(?). The work presented in this paper most closely resembes the topics listed here. Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.

% Contributions (decided to cut for now)
% --------------------------------------

% \subsection{Contributions}

>>>>

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities. The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}

We define the task of automated detection and tracking of arbitrary objects in videos to be three things: extraction, localization, and tracking. Extraction is the task of determining the spatiotemporal regions of a video that constitute objects, localization is the task of finding the positions and/or shapes of distinct objects, and tracking is the task of maintaining the identities of the detected objects over time. An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization. Algorithms able to be applied to a variety of objects and video types are particularly desirable. This paper introduces a new framework involving the use of dependent Dirichlet process mixture models. This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research on methods for general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking. Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus. Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection, localization, and tracking. An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently. Tracking involves finding blobs with similar locations and appearances over time. The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}. These difficulties stem from performing detection via segmentation on individual frames.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method provides both a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and a general framework for identifying and isolating objects during tracking. Specifically, this work describes how a dependent Dirichlet process mixture model can accomplish these goals. The DDP is first reviewed in the abstract. This model can be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then present our specific model for data obtained via a basic extraction procedure. We describe inference algorithms for the DDP, and specifically our specific model, that allow object localization and tracking to be carried out. We demonstrate our implementation on multiple synthetic and benchmark datasets. Standard performance metrics for each supports our hypothesis that general extraction combined withe sophisticated inference can perform detection and tracking at a level comperable to state-of-the-art, object specific algorithms.

% Prior Work
% ----------

\section{Background}
\label{sec:priorwork}

A variety of methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects. These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.

Extraction methods aim to discern the foreground regions of a video from the background. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. Another technique for discerning moving regions in an image involves optical flow \citep{horn_1981, bobick_2001}. In \cite{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are found through analysis of optical flow.

A great deal of research has focused on developing algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.

Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of objects in a video. The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking. This method begins by performing a simple data extraction procedure (that is consistent for all videos), which can be carried out broadly but yields noisy data; we model this noisy extracted data with a sophisticated type of Bayesian mixture model. Inference can then be performed on this model to gain detection and tracking results.

This paper describes the data extraction procedures (Section~\ref{sec:dataextraction}), the general form of the model (Section~\ref{sec:modeldefinition}), the specific distributions used in our implementation of the model (Section~\ref{sec:modelspecification}), inference procedures used to get tracking results (Section~\ref{sec:inference}), and experimental results (Section~\ref{sec:experiments}).

% Overview
% --------

% \section{Overview of Method}
% \label{sec:overviewofmethod}

% Frame Diff and Longimg figure
% -----------------------------

\begin{figure*}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_4.pdf}} 
  %../img/longimg_7_detail.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f). The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
  \label{fig:img_and_framediff}
\end{figure*}

% Data Extraction
% ---------------

\section{Data Extraction}
\label{sec:dataextraction}

A basic set of observations that can be extracted from video are three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel. Extraction procedures, such as those discussed in Section~\ref{sec:introduction}, are used to isolate pixels lying in the regions of a video that represent objects. For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. Accordingly, we carry out a method called ``frame differencing'' for all experiments, which locates regions within video frames that contain motion by recording the positions of pixels exhibiting large enough differences in intensity or value between succesive frames. Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \willie{cite cite cite}). Examples of pixel location data extracted via frame differencing are shown in Fig.~\ref{fig:img_and_framediff}(a-j).

Incorporating color information into our model allows us to infer object color, better maintain the identity of adjacent objects, and track objects through occlusion with greater accuracy. We therefore capture the color of each extracted pixel. Specifically, for each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$. We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of ``color counts''. This vector represents a discrete distribution over colors in the vicinity of a pixel. We will refer to this extraction technique as ``color counting''.

The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index. 
Additional features may be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations. We leave these attempts for future studies.

% We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:priorwork}. The most basic dataset that can be gained from extraction is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region found during extraction and $t$ is the frame in which the pixel resides. Once this dataset, which we will refer to as a ``baseline extraction dataset'', has been found, additional features may also be extracted from the video. Examples of additional features include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, and are applicable to a wide variety of object types.

% The baseline extraction dataset has the appearance of worm-like clusters that follow the paths of individual objects as they move (as illustrated in Figure~\ref{fig:longimg_all}). Modeling this data with a dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpmixture}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data. By choosing this type of model, we aim to provide a clustering result that isolates each worm-like cluster and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inferencetotrackingresults}).

% For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. Consequently, in all experiments we carry out a method called ``frame-differencing'', which locates motion by recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold. Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods that provide accurate foreground segmentation for non-stationary video). We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like clusters from a variety of videos containing multiple moving objects. A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure~\ref{fig:img_and_framediff}(a-i).

% We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model. For each pixel $x = (i, j, t)$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$. We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'. We will refer to this extraction technique as `color counting'.
